---
title: "WhetherWeatherAffectsCrime"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r}
```
```{r}
#Load libraries
library(tidyverse)
library(ggplot2)
library(lubridate)

dataset = read.csv("Crime20161718.csv", stringsAsFactors = F, header = T)
dataset$Date = as_datetime(dataset$Date)
data = read.csv("CrimeWeather2010.csv", stringsAsFactors = F, header = T)
data$Date = as_datetime(mdy(data$Date))
```

### Crime Counts across Years
```{r echo=FALSE}
data %>%
  ggplot(aes(Date,TotalCrimes)) + 
  geom_point(col = "tomato", alpha = 0.3) + 
  geom_smooth(se = F, col="tomato") +
  geom_point(aes(Date,Arrest),col = "royalblue", alpha = 0.3) +
  geom_smooth(aes(Date,Arrest),se = F, col="royalblue") +
  xlab("Year") +
  ylab("Total Crimes/Arrests") +
  ggtitle("Crime Trend Across Years") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5))
```

###Crime Counts Monthly
```{r echo=FALSE}
data %>% filter(year(Date)==2017) %>%
  group_by(month = month(Date,label=T)) %>%
  summarise(Count = sum(TotalCrimes)) %>%
  ggplot(aes(month,Count)) + 
  geom_bar(stat="identity", fill = "royalblue", alpha = 0.7) +
  scale_fill_gradient2(high = "tomato") +
  xlab("Months (2018)") +
  ylab("Count of Crimes") +
  ggtitle("Crime Trend") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "None")
```

###Crime Counts Monthly across Years
```{r echo=FALSE}
data %>% filter(between(year(Date),2010,2017)) %>%
  group_by(Year=year(Date),month = month(Date,label=T)) %>%
  summarise(Count = sum(TotalCrimes)) %>%
  ggplot(aes(month,Count,group=1)) + 
  geom_point(col = "royalblue", alpha = 0.5) + 
  geom_line(col = "royalblue", alpha = 0.5) + 
  facet_wrap(~Year) +
  xlab("Months") +
  ylab("Count of Crimes") +
  ggtitle("Crimes Across Years") +
  theme_bw()
```

###Crime Counts - Day of the Week and Month of the Year
```{r}
data %>% filter(year(Date)==2017) %>%
  group_by(month = month(Date,label=T), 
           weekday = wday(Date,label=TRUE,week_start = 1)) %>%
  summarise(Count = sum(TotalCrimes)) %>% 
  ggplot(aes(x=month,y=weekday,fill=Count)) +
  geom_tile(color = "black") +
  xlab("Month of the Year") +
  ylab("Day of the Week") +
  ggtitle("Crime Counts Across the Year - Daily") +
  scale_fill_gradient(name = "Crime Counts",low = "#F5F5DC",high = "#36648B") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

###Crime Counts - Hour of the day and Month of the Year
```{r}
dataset %>% filter(year(Date)==2017) %>%
  group_by(month = month(Date,label=T), 
           hour = hour(Date)) %>%
  summarise(Count = n()) %>% 
  ggplot(aes(x=month,y=as.factor(hour),fill=Count)) +
  geom_tile(color = "black") +
  xlab("Month of the Year") +
  ylab("Hour of the Day") +
  ggtitle("Crime Counts Across the Year - Hourly") +
  scale_fill_gradient(name = "Crime Counts",low = "#F5F5DC",high = "#36648B") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

###Crime Counts - Hour of the day and Day of the Week
```{r}
dataset %>% filter(year(Date)==2017) %>%
  group_by(weekday = wday(Date,label=TRUE,week_start = 1), 
           hour = hour(Date)) %>%
  summarise(Count = n()) %>% 
  ggplot(aes(x=weekday,y=as.factor(hour),fill=Count)) +
  geom_tile(color = "black") +
  xlab("Month of the Year") +
  ylab("Hour of the Day") +
  ggtitle("Crime Counts Across the Year - Hourly") +
  scale_fill_gradient(name = "Crime Counts",low = "#F5F5DC",high = "#36648B") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

###Create Crime Type Counts
```{r}
dataset %>% filter(Year==2018) %>% 
  group_by(PrimaryType) %>% 
  summarise(Count = n()) %>%
  ggplot(aes(reorder(PrimaryType,Count),Count)) + 
  geom_bar(stat="identity", fill = "royalblue", alpha = 0.7) + 
  coord_flip() +
  xlab("Crime Type") +
  ylab("Number of Crimes") +
  ggtitle("Crimes Based on Crime Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

###Daily crimes - All Crimes
```{r}
data %>% 
  filter(year(Date)==2018) %>% 
  group_by(weekday = wday(Date,label=TRUE,week_start = 1)) %>%
  summarise(Count = sum(TotalCrimes)) %>%
  ggplot(aes(weekday,Count)) + 
  geom_bar(stat="identity", fill = "royalblue", alpha = 0.7) +
  scale_fill_gradient2(high = "tomato") +
  xlab("Months (2018)") +
  ylab("Count of Crimes") +
  ggtitle("Crime Trend") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "None")
```

###Daily Crimes - Crime Type
```{r}
dataset %>% 
  filter(Year==2018,
         PrimaryType %in% c("THEFT","BATTERY","CRIMINAL DAMAGE","ASSAULT","ROBBERY")) %>% 
  group_by(PrimaryType,weekday = wday(Date,label=TRUE,week_start = 1)) %>%
  summarise(Count = n()) %>%
  ggplot(aes(weekday,Count)) + 
  geom_bar(stat="identity", fill = "royalblue", alpha = 0.7) +
  facet_wrap(~PrimaryType) +
  scale_fill_gradient2(high = "tomato") +
  xlab("Months (2018)") +
  ylab("Count of Crimes") +
  ggtitle("Crime Trend") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "None")
```

###Hourly Crime Data
```{r}
dataset %>% 
  filter(Year==2018) %>% group_by(hour = hour(Date)) %>% summarise(Count = n()) %>%
  ggplot(aes(hour,Count)) + 
  geom_bar(stat = "identity", fill="royalblue",alpha=0.7) +
  xlab("Hour of Day") +
  ylab("Count of Crimes") +
  ggtitle("Crime Trend - Hourly") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "None")
```

###Location Based Crimes
```{r}
dataset %>% 
  filter(Year==2018) %>% 
  group_by(LocationDesc) %>% 
  summarise(Count = n()) %>%
  arrange(desc(Count)) %>%
  top_n(25) %>%
  ggplot(aes(reorder(LocationDesc,Count),Count)) + 
  geom_bar(stat = "identity",fill = "royalblue", alpha = 0.7) +
  coord_flip() + 
  xlab("Location") +
  ylab("Count of Crimes") +
  ggtitle("Crimes Based on Location Type (Top 25)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "None")
```

###Temperature Heat Map
```{r}
data %>% filter(year(Date)==2017) %>%
  group_by(month = month(Date,label=T), 
           weekday = wday(Date,label=TRUE,week_start = 1)) %>%
  summarise(Temp = mean(TempAvg)) %>% 
  ggplot(aes(x=month,y=weekday,fill=Temp)) +
  geom_tile(color = "black") +
  xlab("Month of the Year") +
  ylab("Day of the Week") +
  ggtitle("Temperature across the Year") +
  scale_fill_gradient(name = "Crime Counts",low = "#F5F5DC",high = "#36648B") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

##Basic Models
###Linear Regression: Theft
###Creating Dataset
```{r}
TheftData = dataset %>% filter(PrimaryType=="THEFT",Year==2018) %>%
  group_by(date = date(Date)) %>%
  summarise(TotalCrimes = n(), TempAvg = mean(TempAvg), 
            WindAvg = mean(WindAvg), Precipitation = mean(Precipitation),
            Snow = mean(Snow), SnowDepth = mean(SnowDepth), 
            IndFog = mean(IndFog), IndHeavyFog = mean(IndHeavyFog),
            IndThunder = mean(IndThunder), IndPellets = mean(IndPellets),
            IndGlaze = mean(IndGlaze), IndSmoke = mean(IndSmoke),
            IndDriftSnow = mean(IndDriftSnow)) %>%
  mutate_at(vars(IndFog,IndHeavyFog,IndThunder,IndPellets,IndGlaze,IndSmoke,IndDriftSnow), factor)
```
###Building Linear Model
```{r}
lm_theft = lm(TotalCrimes~TempAvg, data = TheftData)
summary(lm_theft)
```

##Checking Assumptions
###Homoscedasity
```{r}
TheftData %>% 
  ggplot(aes(x=1:length(lm_theft$residuals),y=lm_theft$residuals)) + 
  geom_point(col = "royalblue", alpha = 0.7) + 
  geom_smooth(se=F, size=2) +
  xlab("Index") +
  ylab("Model Residuals") +
  ggtitle("Assumption Check: Homoscedasity") +
  theme_bw()
```

###Normality
```{r}
TheftData %>% 
  ggplot(aes(x=lm_theft$residuals)) + 
  geom_density(fill = "royalblue", alpha = 0.5) + 
  xlab("Residuals") +
  ylab("Density") +
  ggtitle("Assumption Check: Normality") +
  theme_bw()
```

###Plotting the linear Regression
```{r}
TheftData %>%
  ggplot(aes(TempAvg,TotalCrimes)) + 
  geom_point(col = "royalblue", alpha = 0.7) +
  geom_line(data=TheftData,aes(x=TempAvg,y=lm_theft$fitted.values), color = "royalblue",size = 2) +
  xlab("Average Temperature (in Farheneit)") +
  ylab("Total Crimes") +
  ggtitle("Effect of Temperature on Crime - THEFT") +
  theme_bw()
```
